Prestige as a Driving Force in Cultural Transmission

### 2022-05-29 ms/PrestigeBias.tex
update

Inspired by a model by \citet{evolutionBook}, we developed a cultural-evolution model with prestige bias, which included both success and influence biases, where the latter is a bias towards role-models with many copiers.
As shown in \cref{fig:influence_advantage}, given the option to choose the ratio of success and influence, copiers will minimize their estimation error, hence will copy the traits that are closer to the ideal phenotype. We see that throughout the choosing process, the estimation error decreases, and the success bias rate also decreases. This suggests that the later a copier gets to choose, the less he should rely on his own estimation, and more on previous choices.
In addition, we see that the process of choosing itself is necessary, and when using a constant ratio ($\alpha$), the mean estimation error in the population is greater by a factor of $10$, even when using the $\alpha$ the population converged to when being able to choose it (\cref{fig:const_influence}).
After we showing that choosing one's success bias weight is beneficial to the copier, we continued to create our models- continuous and dichotomous, that incorporated both success and influence. However, running simulations for a nested iterative model is costly, and analyzing such model mathematically is very hard. We therefor found approximations for the role-model choice process: the generalized binomial distribution (\cref{res:GBD}) and the Dirichlet-Multinomial distribution (\cref{cor:dirichlet}).
To use approximate our model with the GBD distribution, we must assume homogeneous estimation error, $e$, and homogeneous success bias weight, $\alpha$. Both assumptions are necessary to this approximation, because it requires all copiers to be identical (order of copiers must not be a factor).
For the DM approximation we can relax the assumption for $\alpha$, and allow a varying $\alpha_j$ for each role-model $j$. This means that a copier is still not able to choose a success bias weight for itself, and instead it is a characteristic of the role-model only.
Using the DM approximation, because it is easier to analyze, we found equivalents for Kimura's equations, for our dichotomous model.
Following \cref{res:selection_coef,res:effective_population} we found an approximation for the fixation probability of the ideal trait, and an approximation for the time of fixation. These approximations hold only when the equivalent for the selection coefficient, in our model $1-\beta(A)$, is low (0.4 or lower). We also discovered that changing the success bias weight doesn't hurt the goodness of the approximations, and only affects the effective population size $N_e$.
Once we found the equations for a constant environment, we moved on to a cyclic changing environment.
Based on \cref{res:ch_expected,res:ch_var}, we found the approximations for the fixation probability and time to fixation.
We discovered that as in the constant environment, the effective population size doesn't affect the goodness of fit of the approximation. The success coefficient however, must be even lower than before for the approximation to fit. When the success coefficient is higher than 0.15, the simulation results were located above the changing environment approximation, and below the constant environment approximation. We believe the reason is the structure of the cycle.
Our proof and approximation in the changing environment are for a large amount of cycles, and when the success coefficient is too high, there might be very few cycles. Either the ideal trait is copied by enough copiers so that the influence is sufficient to negate the success bias when the cycle changes (and the beneficial trait becomes the maladaptive), or the opposite happens, and the ideal trait gets extinct before there were enough copiers that copied it.
We then tried to change the ratio between the number of cycles where $\hat{A}$ is beneficial and maladaptive. We showed that the approximation fits well regardless of the ratio, but when the ratio of adaptive generations to maladaptive ones is very high, it is very similar to a constant environment model.

@@

@@

\paragraph{Further work.} %TODO revisit
Our model can be expanded in various ways: observing the effects of different bias functions, 
including errors when estimating influence or costs when copying from an influential role-model, and inclusion of multiple, rather than a single, optimal phenotype.

>>>>

Inspired by a model by \citet{evolutionBook}, we developed a cultural-evolution model with prestige bias, which included both success and influence biases, where the latter is a bias towards role-models with many copiers.
As shown in \cref{fig:influence_advantage}, given the option to choose the ratio of success and influence, copiers will minimize their estimation error, hence will copy the traits that are closer to the ideal phenotype. We see that throughout the choosing process, the estimation error decreases, and the success bias rate also decreases. This suggests that the later a copier gets to choose, the less he should rely on his own estimation, and more on previous choices.
This behavior fits the intuitive thought - the more information you have, using others as an information source, the more informative your choice can be, therefore minimizing the distance from your choice of trait to the ideal one.

In addition, we see that the process of choosing itself is necessary, and when using a constant ratio ($\alpha$), the mean estimation error in the population is greater by a factor of $10$, even when using the $\alpha$ the population converged to when being able to choose it (\cref{fig:const_influence}).
After we showing that choosing one's success bias weight is beneficial to the copier, we continued to create our models- continuous and dichotomous, that incorporated both success and influence. However, running simulations for a nested iterative model is costly, and analyzing such model mathematically is very hard. We therefor found approximations for the role-model choice process: the generalized binomial distribution (\cref{res:GBD}) and the Dirichlet-Multinomial distribution (\cref{cor:dirichlet}).
To use approximate our model with the GBD distribution, we must assume homogeneous estimation error, $e$, and homogeneous success bias weight, $\alpha$. Both assumptions are necessary to this approximation, because it requires all copiers to be identical (order of copiers must not be a factor).
For the DM approximation we can relax the assumption for $\alpha$, and allow a varying $\alpha_j$ for each role-model $j$. This means that a copier is still not able to choose a success bias weight for itself, and instead it is a characteristic of the role-model only.
Using the DM approximation, because it is easier to analyze, we found equivalents for Kimura's equations, for our dichotomous model.
Following \cref{res:selection_coef,res:effective_population} we found an approximation for the fixation probability of the ideal trait, and an approximation for the time of fixation. These approximations hold only when the equivalent for the selection coefficient, in our model $1-\beta(A)$, is low (0.4 or lower). We also discovered that changing the success bias weight doesn't hurt the goodness of the approximations, and only affects the effective population size $N_e$.
Once we found the equations for a constant environment, we moved on to a cyclic changing environment.
Based on \cref{res:ch_expected,res:ch_var}, we found the approximations for the fixation probability and time to fixation.
We discovered that as in the constant environment, the effective population size doesn't affect the goodness of fit of the approximation. The success coefficient however, must be even lower than before for the approximation to fit. When the success coefficient is higher than 0.15, the simulation results were located above the changing environment approximation, and below the constant environment approximation. We believe the reason is the structure of the cycle.
Our proof and approximation in the changing environment are for a large amount of cycles, and when the success coefficient is too high, there might be very few cycles. Either the ideal trait is copied by enough copiers so that the influence is sufficient to negate the success bias when the cycle changes (and the beneficial trait becomes the maladaptive), or the opposite happens, and the ideal trait gets extinct before there were enough copiers that copied it.
We then tried to change the ratio between the number of cycles where $\hat{A}$ is beneficial and maladaptive. We showed that the approximation fits well regardless of the ratio, but when the ratio of adaptive generations to maladaptive ones is very high, it is very similar to a constant environment model.

@@

Furthermore, according to \citet{no_replication}, the process of cultural evolution doesn't require replication of representations.
They base their assumption on three claims - mental representations are non-discrete, cultural transmission is highly inaccurate, and mental representations are not replicated, but rather are 'reconstructed' through an inferential process that is strongly affected by cognitive 'attractors.' In their paper they describe three different models to support their points.
We see a high similarity between their model and ours. Like them, we treat the cultural trait as non-discrete (in the main model, before simplifying it to dichotomous in order to analyze it). We also assume error in estimation, which the copiers aren't really replicating their role-models, but rather reconstructing them to create a different trait, which is their representation of the role-model's trait. In addition, the inferential process that they describe as strongly affected by cognitive attractors match our influence bias, resulting in a very similar base assumptions in our models.

@@

\paragraph{Further work.}
In this paper we described several models and processes, analyzed them, and discovered some interesting approximations and results. But even though, there's much more to research in the direction we took, which can help us better model, understand, and maybe even predict, human behavior, specifically in this era where social media is very accessible and informative.
Specifically, a deeper analysis of the first model, where every copier chooses its $\alpha$, is preferred. It would be interesting to see the if the mean error value and the $\alpha$ are converging to specific values, and what they depend on.
Following that, it is possible to relax some assumptions we took in our main model in order to find approximations for it - like homogeneous estimation error and success bias weight.
Lastly, it would be interesting to analyze the continuous model, instead of its dichotomous version. As we've discussed it is probably a much more accurate model of cultural evolution.
